Theory about Normal Basis on $F(2^k)$

### 2017-01-24 dissertation/submission/intro.tex
dissertation after defense update 2
dissertation: after submission_1st revision

\chapter{Introduction}
\label{ch:intro}

\section{Hardware Design and Verification Overview}
During the past decades, the level of integration in modern VLSI systems is becoming higher and higher because of the Moore's law.
As a result, an entire system with billions of transistors can be built upon a single chip.
The design process also evolves from manual design with little validation, to 
a formal 3-step procedure which requires collaboration of teams with large number of 
engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
2) Implement, which is to refine the design through various abstraction levels with the assistance of Computer-Aided-Design (CAD)
tools; 3) Verify, which is to verify the correctness of design and implementation.

@@

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.7\textwidth]{newfig/designflow.eps}
}
\caption{Typical hardware design flow}
\label{fig:designflow}
\end{figure}
}

@@

\section{Formal Verification: Property and Equivalence Checking}

{\it Property checking} (or property verification) verifies
that a design satisfies certain given properties. Property checking is done mainly 
in the form of theorem proving (TP), model checking (MC), or TP/MC hybrid approaches.
\begin{enumerate}[{1)}]
\item \emph{Theorem proving} \cite{theoremproving:91} 
is a method of reasoning and mathematical logic dealing with proving 
mathematical theorems. In the application to circuit property checking, 
the specification as well as the circuit implementation
is described as theorems in mathematical logic. Subsequently logic rules
are employed to deduce new objective theorems. In practice, the tool can reduce
a proof goal to simpler sub-goals for automatic verification.

@@

@@

\begin{figure}[h]
\centering{
\includegraphics[scale=0.3]{newfig/fig_SAT.eps}
\caption{An example of equivalence checking on miter of circuit $A$ and $B$ using SAT}
\label{fig:miter}}
\end{figure}

>>>>

\chapter{Introduction}
\label{ch:intro}
\vspace{-0.8cm}
\section{Hardware Design and Verification Overview}
During the past decades, the level of integration in modern VLSI systems is becoming higher and higher because of the Moore's law.
As a result, an entire system with billions of transistors can be built upon a single chip.
The design process also evolves from manual design with little validation, to 
a formal 3-step procedure which requires collaboration of teams with large number of 
engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
2) Implement, which is to refine the design through various abstraction levels with the assistance of Computer-Aided-Design (CAD)
tools; 3) Verify, which is to verify the correctness of design and implementation.

@@

@@

\section{Formal Verification: Property and Equivalence Checking}
{\it Property checking} (or property verification) verifies
that a design satisfies certain given properties. Property checking is done mainly 
in the form of theorem proving (TP), model checking (MC), or TP/MC hybrid approaches.
\begin{enumerate}[{1)}]
\item \emph{Theorem proving} \cite{theoremproving:91} 
is a method of reasoning and mathematical logic dealing with proving 
mathematical theorems. In the application to circuit property checking, 
the specification as well as the circuit implementation
is described as theorems in mathematical logic. Subsequently logic rules
are employed to deduce new objective theorems. In practice, the tool can reduce
a proof goal to simpler sub-goals for automatic verification.

@@

\begin{figure}[bp]
\centerline{
\includegraphics[width=0.7\textwidth]{newfig/designflow.eps}
}
\caption{Typical hardware design flow.}
\label{fig:designflow}
\end{figure}

@@

\begin{figure}[bp]
\centering{
\includegraphics[scale=0.3]{newfig/fig_SAT.eps}
\caption{An example of equivalence checking on miter of circuit $A$ and $B$ using SAT.}
\label{fig:miter}}
\end{figure}

### 2017-01-24 dissertation/submission/normal.tex
dissertation: after submission_1st revision

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/Moore.eps}
\caption{A typical Moore machine and its state transition graph.}
\label{fig:Moore}}
\end{figure}

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/convention.eps}
\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking.}
\label{fig:convention}}
\end{figure}

@@

\section{Formal Verification of Normal Basis Multipliers using Gr\"obner Bases}
The gate-level design of a NB multiplier can be generated using approaches introduced in 
Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
an approach similar to that in Chapter \ref{ch:reacha}. 
First we introduce the sketch of our approach using abstraction term order (ATO)
mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
the concept of ``RATO", which is previously used in Section \ref{sec:improve}.
% The following part borrows contents
% from my own conference paper \cite{myDATE}.
\subsection{Implicit Unrolling based on Abstraction with ATO}
\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
\caption{Architecture of a combinational GF multiplier.}
\label{fig:comb_RAB}}
\end{figure}

The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
$intermediate~variables~>~R~>~A,B$ for the circuit in Figure \ref{fig:comb_RAB} to compute Gr\"obner basis, 
the function of the combinational logic component can be abstracted as 
$$R = \Func(A,B)$$
To verify the functional correctness of a combinational NB multiplier ({\it e.g.} Mastrovito multiplier
or Montgomery multiplier), the function given by abstraction will be computed as
$$R=A\cdot B$$
While in the sequential case, the function of combinational logic only fulfills a part of the multiplication.
For example, in the RH-SMPO introduced in Section \ref{sec:nbdesign}, 
the combinational logic actually implements  $F_{k-1}$ in Equation \ref{eqn:aux},
while computing all of $F_i$ requires $k$-cycle unrolling of the circuit.
Nevertheless, the abstraction still provides a word-level 
representation which can be more efficient for unrolling than bit-level expressions.
In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
explicit unrolling and avoid bit-blasting problem.

For 2-input sequential NB multipliers, abstraction is utilized to implement following algorithm:
%\vspace{-0.1in}
\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Circuit polynomial ideal $J_{ckt}$, vanishing ideal $J_0$, initial
   state ideal for output $R (=0)$, inputs $\mathcal{G}(A_{init}), \mathcal{H}(B_{init})$} 
\KwOut{Circuit function $R_{final}$ after $k$ cycles of $A_{init}, B_{init}$}
  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  \tcc{$A, B$ are word-level variables, solutions to polynomial equations 
  $\mathcal{G},\mathcal{H}$ denote the initial values of preloaded operands $A_{init},B_{init}$}
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J_{ckt} + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	\tcc{Using projections of varieties from abstraction theory, we get NS variables $R',A',B'$
	in terms of PS $A,B$}
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
	\tcc{By replacing NS variables with PS variables we push it to next time-frame}
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
\DecMargin{1em}

Notations in the following figure as well as in next example comply with 
the notations in this algorithm.

\begin{figure}[H]
\centering{
\includegraphics[width=5in]{newfig/gf_seq_model.eps}
\caption{A typical normal basis GF sequential circuit model. $A =
  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
  $A', B', R'$ denote next-state inputs.}
\label{fig:sequential}}
\end{figure}

@@

In the practice of sequential GF multiplier verification, the combinational 
logic actually implements function not only related to current operands $A$ and $B$,
but also involved with PS variable ({\it i.e.} temporary product) $R$. Which can be 
obtained using the abstraction:
$$R' = \Func(A,B,R)$$

Using ATO, if we put $R$ ahead of $R',A,B$ in term ordering, $R$ is thus eliminated
and the result in single iteration will be $R' = \Func(A,B)$. NS operands 
$A'$ and $B'$ are right-cyclic shift of $A$ and $B$, which can be directly written.
If initial values $A_{init}, B_{init}$ are treated as parameters, the NS ideal 
contains polynomials with $A', B'$ and $R' = \Func(A_{init},B_{init})$. This 
is also shown in the following example.

\begin{Example}[Functional verification of 5-bit RH-SMPO]
\label{ex:RHSMPO}
Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO, with normal element $\beta = \alpha^5$. The transition function for
operands $A,B$ perform cyclic shift, while the transition function for $R$ has to be computed through the Gr\"obner basis
abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
all gates in combinational logic block and definition of word-level variables.
$$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$

\begin{eqnarray*}
 \left. 
\begin{aligned}
d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4\\
c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2 \\
R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4
\end{aligned}\right\} ~~~\{f_1\dots f_{15}\}\\
 \left.
\begin{aligned}
 A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}\\
		  B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\\
		  R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18}\\
		  R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18}
		 \end{aligned}\right\} ~~~\{f_{16}\dots f_{19}\}
\end{eqnarray*}

@@

\subsection{Overcome Computational Complexity using RATO}
We implemented Algorithm \ref{alg:modified} in \textsc{Singular}, with similar setup and environments as
with the experiments in Section \ref{sec:exp_reacha}. The result of verifying Agnew's SMPO is shown 
in following table:

\begin{table}[H]
\begin{center}
	    \caption{Runtime of Gr\"obner Basis Computation of Agnew's SMPO in Singular using ATO $>$.}\label{tab:slimgb}
	    \begin{tabular}{|c|c||c|} 
	        \hline
		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (seconds)   \\
		\hline
	        $5$	&  $65$  & $16$ \\
	        $6$ &  $90$  & $368$ \\
	        $9$ &  $189$ & $43602$ \\
		$11$	&  $275$  & $TO$ ($>48$ hr)  \\
	        \hline
	    \end{tabular}
	\end{center} 
\end{table}

@@

\subsection{The Overall Verification Approach}
Based on the above concepts and improvements, the functional verification for sequential GF multiplier 
on word-level with $k$-bit input operands $A,B$ and $k$-bit output $R$ is described as follows:
\begin{enumerate}[{1)}]
\item Given a sequential GF multiplier $S$, with word-level $k$-bit inputs $A,B$ and output $R$.
\item Given a primitive polynomial $P(x)$ of degree $k$ and construct $\Fkk$, and let $P(\alpha)=0$.
\item Perform a reverse-topological traversal of $S$ to derive RATO: LEX with $\{x_1>x_2>\cdots>x_d>R'>R>A>B\}$,
where $\{x_1,\dots,x_d\}$ are bit-level variables of $S$, $R'$ is the NS output value, and 
$R,A,B$ are word-level PS variables.
\item Derive the set of bit-level polynomials $\{f_1,\dots,f_s\}$ from each gate in $S$, and represent them 
using RATO. These polynomials are in the form $x_i+tail(f_i)$ where $x_i$ is the output of the corresponding logic gate.
\item Compose the bit-level to word-level correspondences polynomials such as $r_0\beta+r_1\beta^2+\cdots+r_{k-1}\beta^{2^{k-1}}+R$.
Add them to set $\{f_1,\dots,f_s\}$ and generate ideal $J_{ckt}$. Compose the ideal of vanishing polynomials
$J_0 = \langle x_i^2+x_i, \dots, R^{2^k}+R,\dots\rangle$.
\item Select the only critical pair $(f_w,f_g)$ that does not have relatively prime leading terms. Compute
$Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}r$.
\item Construct matrices $M_0,\dots,M_{k-1}$, where $M_i$ is $M$ with the $i$-column replaced by vector 
$[R~~R^2~~\cdots~~R^{2^{k-1}}]^T$, and $M$ is the reverse-circulant matrix generated by vector $(\beta,\beta^2,\dots,\beta^{2^{k-1}})$.
\item Symbolically compute the determinants $|M_i|$ to find set $F_R$, where $f_{R_i}:R_i+|M_i|$, for $0\leq i \leq k-1$.
Obtain $F_A, F_B$ from $F_R$ since they have the same form.
\item Compute $r \xrightarrow{F_A\cup F_B \cup F_R}_{+}r_w$, and iterate $r_w$ as $to^i$ in Algorithm \ref{alg:modified}.
{\bf Then $r_w$ after iteration $k$ is of the form $R'+\Func(A,B)$ and it is the unique, canonical word-level abstraction
of $S$ over $\Fkk$ after $k$ clock-cycles}.
\end{enumerate}

@@

\vspace{-0.1in}
\begin{table}[htb]
\centering
\caption{\small Run-time for verification of bug-free RH-SMPO circuits
  for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs}
\label{tbl:equiv}
\begin{tabular}{|c||c|c|c|c|} 
\hline
& \multicolumn{4}{|c|}{Word size of the operands $k$-bits}  \\
\hline
Solver & 11 & 18 & 23 & 33 \\
\hline
\hline
Lingeling & 593  & \emph{TO}  & \emph{TO}  & \emph{TO}\\
\hline
\hline
ABC & 6.24 & \emph{TO} & \emph{TO} & \emph{TO}\\
\hline
\hline
BDD & 0.1 & 11.7 & 1002.4 & \emph{TO}  \\
\hline
\end{tabular}
\label{table:satbdd}  
\end{table}

@@

\begin{table}[htb]
\centering
\caption{\small Similarity between RH-SMPO and Agnew's SMPO}
\label{tbl:fraig}
\begin{tabular}{|c||c|c|c|c|} 
\hline
Size $k$ & 11 & 18 & 23 & 33 \\
\hline
$N_1$ & 734  & 2011  & 3285  & 6723\\
\hline
$N_2$ & 529 & 1450 & 2347 & 4852\\
\hline
Similarity & $27.9\%$ & $27.9\%$ & $28.6\%$ & $27.8\%$  \\
\hline
\end{tabular}\label{table:similarity}  
\end{table}

@@

\begin{table}[htb]
\centering
\caption{\small Run-time (seconds) for verification of bug-free and
  buggy RH-SMPO with \textsc{Singular}} 
\label{tbl:exp1}  
\begin{tabular}{|c||c|c|c|c|c|c|} 
\hline
Operand size $k$ & 33 & 51 & 65 & 81 & 89 & 99\\
\hline
\#variables & 4785 & 11424 & 18265 & 28512 & 34354 & 42372\\
\hline
\#polynomials & 3630 & 8721 & 13910 & 21789 & 26255 & 32373\\
\hline
\#terms & 13629 & 32793 & 52845 & 82539 & 99591 & 122958\\
\hline
\hline
Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 & 67021\\
\hline
Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120 & 66929\\
\hline
\end{tabular}
%\begin{minipage}{8cm}
%\def\arraystretch{1.5}\tabcolsep 2pt
%% \begin{tabular}{|c||c|c|c|c|c|} 
%% \hline
%% Operand size $k$ & 33 & 51 & 65 & 81 & 89 \\
%% \hline
%% \#variables & 4785 & 11424 & 18265 & 28512 & 34354\\
%% \hline
%% \#polynomials & 3630 & 8721 & 13910 & 21789 & 26255\\
%% \hline
%% \#terms & 13629 & 32793 & 52845 & 82539 & 99591 \\
%% \hline
%% %\hline
%% Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 \\
%% \hline
%% Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120\\
%% \hline
%% \hline
%% \end{tabular}
%% %\end{minipage}
\end{table}

\begin{table}[hbt]
\centering
\caption{\small Run-time (seconds) for verification of
  bug-free and buggy Agnew's SMPO with \textsc{Singular}}
\label{tbl:exp2}
\begin{tabular}{|c||c|c|c|c|c|} 
\hline
Operand size $k$ & 36 & 66 & 82 & 89 & 100 \\
\hline
\#variables & 6588 & 21978 & 33866 & 39872 & 50300\\
\hline
\#polynomials & 2700 & 8910 & 13694 & 16109 & 20300\\
\hline
\#terms & 12996 & 43626 & 67322 & 79299 & 100100 \\
\hline
\hline
Runtime(bug-free) & 113 & 3673 & 15117 & 28986 & 50692 \\
\hline
Runtime(buggy) & 118 & 4320 & 15226 & 31571 & 58861\\
\hline
\end{tabular}
\end{table}

\begin{table}[hbt]
\centering
\caption{\small Run-time (seconds) for verification of
  RH-SMPO and Agnew's SMPO using C++ implementation}
\label{tab:Cpp}
{\small 
\begin{tabular}{|c|c||c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c||}{\centering Operand size $k$} & 36 & 60 & 81 & 100 & 131 & 162 \\
\hline
\multirow{2}{2cm}{\centering RH-SMPO} & \#Polynomials & 4716 & 12960 & 21870 & 35600 & 56592 & 92826 \\
\cline{2-8}
 & Runtime & 14.3 & 213.3 & 1343 & 4685 & 26314 & 124194 \\
\hline
\multirow{2}{2cm}{\centering Agnew's SMPO} & \#Polynomials & 2700 & 7380 & 13356 & 20300 & 34715 & 52974 \\
\cline{2-8}
 & Runtime & 10.2 & 212.0 & 2684 & 4686 & 56568 & 119441 \\

>>>>

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/Moore.eps}
\caption{A typical Moore machine and its state transition graph.}
\label{fig:Moore}}
\end{figure}

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/convention.eps}
\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking.}
\label{fig:convention}}
\end{figure}

@@

\section{Formal Verification of Normal Basis Multipliers using Gr\"obner Bases}
The gate-level design of a NB multiplier can be generated using approaches introduced in 
Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
an approach similar to that in Chapter \ref{ch:reacha}. 
First we introduce the sketch of our approach using abstraction term order (ATO)
mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
the concept of ``RATO", which is previously used in Section \ref{sec:improve}.
% The following part borrows contents
% from my own conference paper \cite{myDATE}.
\subsection{Implicit Unrolling based on Abstraction with ATO}
The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
$intermediate~variables~>~R~>~A,B$ for the circuit in Figure \ref{fig:comb_RAB} to compute Gr\"obner basis, 
the function of the combinational logic component can be abstracted as 
$$R = \Func(A,B)$$
To verify the functional correctness of a combinational NB multiplier ({\it e.g.} Mastrovito multiplier
or Montgomery multiplier), the function given by abstraction will be computed as
$$R=A\cdot B$$
While in the sequential case, the function of combinational logic only fulfills a part of the multiplication.
For example, in the RH-SMPO introduced in Section \ref{sec:nbdesign}, 
the combinational logic actually implements  $F_{k-1}$ in Equation \ref{eqn:aux},
while computing all of $F_i$ requires $k$-cycle unrolling of the circuit.
Nevertheless, the abstraction still provides a word-level 
representation which can be more efficient for unrolling than bit-level expressions.
In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
explicit unrolling and avoid bit-blasting problem.

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
\caption{Architecture of a combinational GF multiplier.}
\label{fig:comb_RAB}}
\end{figure}

%\vspace{-0.1in}
% \IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Circuit polynomial ideal $J_{ckt}$, vanishing ideal $J_0$, initial
   state ideal for output $R (=0)$, inputs $\mathcal{G}(A_{init}), \mathcal{H}(B_{init})$} 
\KwOut{Circuit function $R_{final}$ after $k$ cycles of $A_{init}, B_{init}$}
  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  \tcc{$A, B$ are word-level variables, solutions to polynomial equations 
  $\mathcal{G},\mathcal{H}$ denote the initial values of preloaded operands $A_{init},B_{init}$}
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J_{ckt} + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	\tcc{Using projections of varieties from abstraction theory, we get NS variables $R',A',B'$
	in terms of PS $A,B$}
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
	\tcc{By replacing NS variables with PS variables we push it to next time-frame}
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
% \DecMargin{1em}

For 2-input sequential NB multipliers, abstraction is utilized to implement Algorithm \ref{alg:modified}.
Notations in Figure \ref{fig:sequential} as well as in next example comply with 
the notations in this algorithm.

\begin{figure}[bp]
\centering{
\includegraphics[width=5in]{newfig/gf_seq_model.eps}
\caption{A typical normal basis GF sequential circuit model. $A =
  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
  $A', B', R'$ denote next-state inputs.}
\label{fig:sequential}}
\end{figure}

@@

In the practice of sequential GF multiplier verification, the combinational 
logic actually implements function not only related to current operands $A$ and $B$,
but also involved with PS variable ({\it i.e.} temporary product) $R$. Which can be 
obtained using the abstraction:
$$R' = \Func(A,B,R)$$
Using ATO, if we put $R$ ahead of $R',A,B$ in term ordering, $R$ is thus eliminated
and the result in single iteration will be $R' = \Func(A,B)$. NS operands 
$A'$ and $B'$ are right-cyclic shift of $A$ and $B$, which can be directly written.
If initial values $A_{init}, B_{init}$ are treated as parameters, the NS ideal 
contains polynomials with $A', B'$ and $R' = \Func(A_{init},B_{init})$. This 
is also shown in the following example.

\begin{Example}[Functional verification of 5-bit RH-SMPO]
\label{ex:RHSMPO}
Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO, with normal element $\beta = \alpha^5$. The transition function for
operands $A,B$ perform cyclic shift, while the transition function for $R$ has to be computed through the Gr\"obner basis
abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
all gates in combinational logic block and definition of word-level variables.
$$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$
\begin{eqnarray*}
 \left. 
\begin{aligned}
d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4\\
c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2 \\
R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4
\end{aligned}\right\} ~~~\{f_1\dots f_{15}\}\\
 \left.
\begin{aligned}
 A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}\\
		  B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\\
		  R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18}\\
		  R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18}
		 \end{aligned}\right\} ~~~\{f_{16}\dots f_{19}\}
\end{eqnarray*}

@@

\subsection{Overcome Computational Complexity using RATO}
We implemented Algorithm \ref{alg:modified} in \textsc{Singular}, with similar setup and environments as
with the experiments in Section \ref{sec:exp_reacha}. The result of verifying Agnew's SMPO is shown 
in Table \ref{tab:slimgb}.

\begin{table}[bp]
\begin{center}
	    \caption{Runtime of Gr\"obner Basis Computation of Agnew's SMPO in Singular using ATO $>$.}\label{tab:slimgb}
	    \begin{tabular}{|c|c||c|} 
	        \hline
		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (seconds)   \\
		\hline
	        $5$	&  $65$  & $16$ \\
	        $6$ &  $90$  & $368$ \\
	        $9$ &  $189$ & $43602$ \\
		$11$	&  $275$  & $TO$ ($>48$ hr)  \\
	        \hline
	    \end{tabular}
	\end{center} 
\end{table}

@@

\subsection{The Overall Verification Approach}
Based on the above concepts and improvements, the functional verification for sequential GF multiplier 
on word-level with $k$-bit input operands $A,B$ and $k$-bit output $R$ is described as follows:
\begin{enumerate}[{1)}]
\item Given a sequential GF multiplier $S$, with word-level $k$-bit inputs $A,B$ and output $R$.
\item Given a primitive polynomial $P(x)$ of degree $k$ and construct $\Fkk$, and let $P(\alpha)=0$.
\item Perform a reverse-topological traversal of $S$ to derive RATO: LEX with $\{x_1>x_2>\cdots>x_d>R'>R>A>B\}$,
where $\{x_1,\dots,x_d\}$ are bit-level variables of $S$, $R'$ is the NS output value, and 
$R,A,B$ are word-level PS variables.
\item Derive the set of bit-level polynomials $\{f_1,\dots,f_s\}$ from each gate in $S$, and represent them 
using RATO. These polynomials are in the form $x_i+tail(f_i)$ where $x_i$ is the output of the corresponding logic gate.
\item Compose the bit-level to word-level correspondences polynomials such as $r_0\beta+r_1\beta^2+\cdots+r_{k-1}\beta^{2^{k-1}}+R$.
Add them to set $\{f_1,\dots,f_s\}$ and generate ideal $J_{ckt}$. Compose the ideal of vanishing polynomials
$J_0 = \langle x_i^2+x_i, \dots, R^{2^k}+R,\dots\rangle$.
\item Select the only critical pair $(f_w,f_g)$ that does not have relatively prime leading terms. Compute
$Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}r$.
\item Construct matrices $M_0,\dots,M_{k-1}$, where $M_i$ is $M$ with the $i$-column replaced by circulant vector 
$[R~~R^2~~\cdots~~R^{2^{k-1}}]^T$, and $M$ is the reverse-circulant matrix generated by circulant vector $(\beta,\beta^2,\dots,\beta^{2^{k-1}})$.
\item Symbolically compute the determinants $|M_i|$ to find set $F_R$, where $f_{R_i}:R_i+|M_i|$, for $0\leq i \leq k-1$.
Obtain $F_A, F_B$ from $F_R$ since they have the same form.
\item Compute $r \xrightarrow{F_A\cup F_B \cup F_R}_{+}r_w$, and iterate $r_w$ as $to^i$ in Algorithm \ref{alg:modified}.
{\bf Then $r_w$ after iteration $k$ is of the form $R'+\Func(A,B)$ and it is the unique, canonical word-level abstraction
of $S$ over $\Fkk$ after $k$ clock-cycles}.
\end{enumerate}

@@

% \vspace{-0.1in}
\begin{table}[bp]
\centering
\caption{\small Run-time for verification of bug-free RH-SMPO circuits
  for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs}
\label{tbl:equiv}
\begin{tabular}{|c||c|c|c|c|} 
\hline
& \multicolumn{4}{|c|}{Word size of the operands $k$-bits}  \\
\hline
Solver & 11 & 18 & 23 & 33 \\
\hline
\hline
Lingeling & 593  & \emph{TO}  & \emph{TO}  & \emph{TO}\\
\hline
\hline
ABC & 6.24 & \emph{TO} & \emph{TO} & \emph{TO}\\
\hline
\hline
BDD & 0.1 & 11.7 & 1002.4 & \emph{TO}  \\
\hline
\end{tabular}
\label{table:satbdd}  
\end{table}

@@

\begin{table}[bp]
\centering
\caption{\small Similarity between RH-SMPO and Agnew's SMPO}
\label{tbl:fraig}
\begin{tabular}{|c||c|c|c|c|} 
\hline
Size $k$ & 11 & 18 & 23 & 33 \\
\hline
$N_1$ & 734  & 2011  & 3285  & 6723\\
\hline
$N_2$ & 529 & 1450 & 2347 & 4852\\
\hline
Similarity & $27.9\%$ & $27.9\%$ & $28.6\%$ & $27.8\%$  \\
\hline
\end{tabular}\label{table:similarity}  
\end{table}

@@

\begin{table}[bp]
\centering
\caption{\small Run-time (seconds) for verification of bug-free and
  buggy RH-SMPO with \textsc{Singular}} 
\label{tbl:exp1}  
\begin{tabular}{|c||c|c|c|c|c|c|} 
\hline
Operand size $k$ & 33 & 51 & 65 & 81 & 89 & 99\\
\hline
\#variables & 4785 & 11424 & 18265 & 28512 & 34354 & 42372\\
\hline
\#polynomials & 3630 & 8721 & 13910 & 21789 & 26255 & 32373\\
\hline
\#terms & 13629 & 32793 & 52845 & 82539 & 99591 & 122958\\
\hline
\hline
Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 & 67021\\
\hline
Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120 & 66929\\
\hline
\end{tabular}
%\begin{minipage}{8cm}
%\def\arraystretch{1.5}\tabcolsep 2pt
%% \begin{tabular}{|c||c|c|c|c|c|} 
%% \hline
%% Operand size $k$ & 33 & 51 & 65 & 81 & 89 \\
%% \hline
%% \#variables & 4785 & 11424 & 18265 & 28512 & 34354\\
%% \hline
%% \#polynomials & 3630 & 8721 & 13910 & 21789 & 26255\\
%% \hline
%% \#terms & 13629 & 32793 & 52845 & 82539 & 99591 \\
%% \hline
%% %\hline
%% Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 \\
%% \hline
%% Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120\\
%% \hline
%% \hline
%% \end{tabular}
%% %\end{minipage}
\end{table}

\begin{table}[bp]
\centering
\caption{\small Run-time (seconds) for verification of
  bug-free and buggy Agnew's SMPO with \textsc{Singular}}
\label{tbl:exp2}
\begin{tabular}{|c||c|c|c|c|c|} 
\hline
Operand size $k$ & 36 & 66 & 82 & 89 & 100 \\
\hline
\#variables & 6588 & 21978 & 33866 & 39872 & 50300\\
\hline
\#polynomials & 2700 & 8910 & 13694 & 16109 & 20300\\
\hline
\#terms & 12996 & 43626 & 67322 & 79299 & 100100 \\
\hline
\hline
Runtime(bug-free) & 113 & 3673 & 15117 & 28986 & 50692 \\
\hline
Runtime(buggy) & 118 & 4320 & 15226 & 31571 & 58861\\
\hline
\end{tabular}
\end{table}

\begin{table}[bp]
\centering
\caption{\small Run-time (seconds) for verification of
  RH-SMPO and Agnew's SMPO using C++ implementation}
\label{tab:Cpp}
{\small 
\begin{tabular}{|c|c||c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c||}{\centering Operand size $k$} & 36 & 60 & 81 & 100 & 131 & 162 \\
\hline
\multirow{2}{2cm}{\centering RH-SMPO} & \#Polynomials & 4716 & 12960 & 21870 & 35600 & 56592 & 92826 \\
\cline{2-8}
 & Runtime & 14.3 & 213.3 & 1343 & 4685 & 26314 & 124194 \\
\hline
\multirow{2}{2cm}{\centering Agnew's SMPO} & \#Polynomials & 2700 & 7380 & 13356 & 20300 & 34715 & 52974 \\
\cline{2-8}
 & Runtime & 10.2 & 212.0 & 2684 & 4686 & 56568 & 119441 \\
